---
title: "Practice"
author: "Marcus Nunes"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: bookdown::pdf_document2
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(GGally)
```

# Instructions

Save a version of this file and call it `practice_answers.Rmd`. Remove this first section and practice what we learned today, writing a small report analyzing the `iris` dataset. 

After your work is done, change the code `output: bookdown::html_document2` on line 5 to `output: bookdown::pdf_document2` in order to have a pdf. It is preferable to work with the HTML version of the report first to save time and create the pdf only after the analysis is complete.



# Introduction

The Iris flower dataset is a well known multivariate dataset introduced in @Fisher1936. It contains 150 observations of three species of plants called _Iris setosa_, _Iris versicolor_, and _Iris virginica_. In this report we will analyze this dataset and show why it is interesting.

# Descriptive Statistics

We will describe the most important features about the Iris flower dataset in this section.

## Tables

1. Create a table with the means for `ìris` dataset. Describe your findings.

```{r q1}
results <- iris %>%
  select_if(is.numeric) %>%
  summarise_all(list(mean = mean))

kable(results, 
      format = "latex", 
      booktabs = TRUE,
      caption = "Regression model result when applied to data `iris`.", 
      digits = 4) %>%
  kable_styling(position = "center")
```

The variable with the highest mean is Sepal.Length, with `r round(results[1], digits = 2)`cm.


## Plots

2. Make scatter plots showing the relationships between all the numeric variables in this dataset. What you can see?

```{r q2, fig.cap="Exploratory data analysis for the iris dataset."}
iris %>%
  select_if(is.numeric) %>%
  ggpairs()
```

It seems all the variables have linear relationships between them. Some are higher and some are lower, but they do seem linear, as we can see in Figure \@ref(fig:q2).

3. Plot the scatter plot between the two variables with the highest correlation, coloring the points according to the flower species. Is there anything special about this plot?

```{r q3, fig.cap="Scatter plot of iris dataset."}
ggplot(iris, aes(x = Petal.Length, y = Petal.Width, colour = Species)) +
  geom_point() +
  labs(x = "Petal Length", y = "Petal Width")
```

\@ref(fig:q3) shows that there are two main clusters in this dataset. One, to the left, with only _Iris setosa_, and other to the right, with both _Iris versicolor_ and _virginica_.


# Modeling

4. Fit a linear regression model between the two variables you found on question 3. Use the variable with the highest mean as the predictor variable.

```{r q4}
fit <- lm(Petal.Width ~ Petal.Length, data = iris)
kable(summary(fit)$coefficients, 
      format = "latex", 
      booktabs = TRUE,
      caption = "Regression model result when applied to data `iris`.", 
      digits = 4) %>%
  kable_styling(position = "center")
```

5. Plot the regression line on the scatter plot.

```{r q5, fig.cap="Linear regression model fitted to iris dataset."}
ggplot(iris, aes(x = Petal.Length, y = Petal.Width)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, colour = "black") +
  labs(x = "Petal Length", y = "Petal Width")
```

6. Make a boxplot comparing the observations of the `Sepal.Width` variable between the three plant species. Do you think there is a group whose mean is different from the others?

```{r q6, fig.cap="Boxplot comparing the means of sepal width for the iris dataset."}
ggplot(iris, aes(x=Species, y=Sepal.Width)) +
  geom_boxplot() +
  labs(x="Species", y="Sepal Width")
```

According to Figure \@ref(fig:q6), it seems _Iris setosa_ has a higher average than the other two species.

7. Test the hypothesis

\begin{align*}
H_0 &: \mu_1 = \mu_2 = \mu_3 \\
H_0 &: \mbox{at least one pair $\mu_i \neq \mu_j$, if $i \neq j$}
\end{align*}

\noindent where $\mu_i$ is the mean for the variable `Sepal.Width` for the groups 

* $i = 1 \mbox{ (setosa})$

* $i = 2 \mbox{ (versicolor})$

* $i = 3 \mbox{ (virginica})$

What is your conclusion?

```{r q7}
fit.anova <- aov(Sepal.Width ~ Species, data = iris)
results.anova <- summary(fit.anova)
options(knitr.kable.NA = "")
kable(results.anova[[1]], 
      format = "latex", 
      booktabs = TRUE,
      caption = "ANOVA results.", 
      digits = 4) %>%
  kable_styling(position = "center")
```

According to Table \@ref(tab:q7), p-value = `r round(results.anova[[1]][1, 5], digits = 4)`. Therefore, we reject $H_0$ at level $\alpha = 0,05$. Therefore, there is at least one pair $\mu_i \neq \mu_j$, if $i \neq j$.


# Conclusion

As we could see, this dataset is great to practice statistical concepts.

# References





